Interpretable Casual Interventions \\ Discourage Spurious Learning

### 2025-03-24 paper/main.tex
added section on sliced wasserstein loss, reformulated ICRM as a constrained optimization problem
refactored contrastive cams as a custom variable to clean-up tex

@@

Per \cite{draelos2020use}, a HiResCAM is defined as follows:
\begin{gather}
	\tilde{\mathcal{\bm{A}}}_c^{\text{HiResCAM}} = \sum^F_{f=1} \frac{\partial s_c}{\partial \bm{A}^f} \odot \bm{A}^f
\end{gather}

Where $s_c$ denotes the logit for a single class. $\tilde{\mathcal{\bm{A}}}_c^{\text{HiResCAM}}$ generates a class-specific CAM. For the CAM architecture, \cite{draelos2020use} (section 3.6.2) shows that:
\begin{gather}
	\sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} = \frac{\bm{w}_c \bm{A}_F}{D_1 D_2} \\
	s_c = \frac{\bm{w}_c \bm{A}_F}{D_1 D_2} + b_c = \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} + b_c
\end{gather}
Therefore, we can recreate the entire input-dependent part of the score using HiResCAMs. Doing this for each class $c$, we obtain the bias-removed logit vector $\in \mathbb{R}^{c}$.

@@

\begin{theorem}\label{classprobdiff}
	Individual class probabilites $\sigma_c(\bm{s})$ depend on the difference(s) between class logits.
\end{theorem}
\begin{proof} Individual class probabilities for logit vector $\bm{s}$ are defined as:
	\begin{gather}
		\sigma_c(\bm{s}) = \frac{e^{s_c}}{\sum_i e^{s_i}}
	\end{gather}
	We define our logit vector in terms of the elementwise difference to a target class $c$:
	\begin{gather}
		\bm{d} = [s_i - s_c]_i \\
		\bm{s} = [s_c + d_i]_i
	\end{gather}
	Based on this definition, class probabilities can now be computed as:
	\begin{align}
		\sigma_c(\bm{s}) &= \frac{e^{s_c}}{\sum_i e^{s_c+d_i}} \\
				 &= \frac{e^{s_c}}{e^{s_c} \sum_i e^{d_i}} = \frac{1}{\sum_i e^{d_i}}
	\end{align}
	This recontextualizes softmax as a direct function of the differences of class logits.
\end{proof}
\begin{corollary} Since class probabilities depend on logit differences, the class probabilities are identitical for any two pairs of logit vectors with identical logit differences.
	\begin{gather}
		\frac{e^{s_c}}{\sum_i e^{s_i}} = \frac{e^{s_c+r}}{\sum_i e^{s_i+r}} \forall r \in \mathbb{R}
	\end{gather}
\end{corollary}
\begin{proof} Recontextualizing $\bm{s}$ per the previous proposition, we get:
	\begin{align}
		\frac{e^{s_c}}{\sum_i e^{s_c + d_i}} &= \frac{e^{s_c+r}}{\sum_i e^{s_c+d_i+r}} \\
		\frac{e^{s_c}}{e^{s_c}\sum_i e^{d_i}} &= \frac{e^{s_c+r}}{e^{s_c + r} \sum_i e^{d_i}} \\
		\frac{1}{\sum_i e^{d_i}} &= \frac{1}{\sum_i e^{d_i}}
	\end{align}
	Therefore, the absolute scale of a logit does not influence the final class probability.
\end{proof}

@@

We therefore derive an optimization objective that attempts to maximize the contrast between the two activations maps. For the binary classification case, $(c, c')$ reflects that correct and incorrect class respectively.
\begin{gather}
	\max_{\theta} s_c - s_{c'} \equiv \max_{\theta} \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} + b_c - \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c',i,j}^{\text{HiResCAM}} - b_{c'}
\end{gather}
Ignoring the bias terms, the new optimization objective becomes:
\begin{gather}
	\max_{\theta} \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, c'), i, j} ~~\text{ where }~~ \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, c')} := \tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}}
\end{gather}
This is equivalent to the original ERM optimization objective, with the difference being the resultant values preserving spatial information in the form of the activation map instead of flattening the feature map. If $\tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}} > 0$, the classification is correct -- if not, the classification is incorrect.

@@

\begin{definition}
	Given a set of classes $C$ with $c_t$ being the target class and $c_{t'}$ being a sampled non-target class, we define ContrastiveCAMs as follows:
	\begin{gather}
		\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c_t, c_{t'})} := \left\{\tilde{\mathcal{\bm{A}}}_{c_t}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c_{t'}}^{\text{HiResCAM}}\right\}^{|C|-1}_{c_{t'} \in c \setminus c_t}
	\end{gather}
\end{definition}

@@

To ensure that the model accurately predicts the target class, each contrastive comparison should yield a positive result. Therefore, the multiclass optimization objective under this construction reads:
\begin{gather}
	\max_{\theta} \sum^{D_1,D_2}_{j,k}\tilde{\mathcal{\bm{A}}}_{(c, c'),j,k}^{\text{contrastive}}\ \forall c' \in \mathbb{Z}_{+}(|C| - 1)
\end{gather}

@@

\begin{theorem}\label{ce-spur-corr}
	Cross-Entropy motivates learning from spurious correlations.
\end{theorem}
\begin{proof} Using the definition of HiResCAMs, ContrastiveCAMs and the elementwise difference setup from Theorem \ref{classprobdiff}, we have:
	\begin{align}
		d &= [s_i - s_c]_i \\
		&= \left[\sum^{D_1,D_2}_{j,k} \tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - \sum^{D_1,D_2}_{j,k} \tilde{\mathcal{\bm{A}}}_{c,j,k}^{\text{HiResCAM}}\right]_i \\
		&= \left[ -\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} \right]_i
	\end{align}
	As an aside: we note that when $i = c$, the contrastive map obtained is $\bm{0} \in \mathbb{R}^{D_1 \times D_2}$, and instead of $|C| - 1$ contrastive maps we now obtain $|C|$ contrastive maps.

	Using our result from Theorem~\ref{classprobdiff}, we have:
	\begin{gather}
		\sigma_c(\bm{s}) = \frac{1}{\sum_i e^{d_i}} = \frac{1}{\sum_i \text{exp}({-\sum \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}})} = \sum_i \text{exp}\left({\sum \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}\right)
	\end{gather}
	For a target class $c$, cross-entropy loss is defined as:
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \sigma_c(\bm{s}) \right) = -\log \left( \sum_i \text{exp}\left({\sum \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}\right) \right)
	\end{gather}
	Vectorizing parameters $\theta \in \mathbb{R}^P$ as well as the contrastive attribution maps $\tilde{\mathcal{\bm{A}}}_{(c, i)}^{\text{contrastive}} \in \mathbb{R}^{D_1 D_2}$, we can use the above result to compute gradient contributions from each pixel within the attribution map:
	\begin{align}
		\nabla_\theta H &= \underbrace{\begin{bmatrix} \frac{\partial H}{\partial \sum \tilde{\mathcal{\bm{A}}}_{(c, i)}^{\text{contrastive}}} \end{bmatrix}_i}_{\mathbb{R}^{1 \times |C|}}
			\cdot \underbrace{\begin{bmatrix} \begin{bmatrix} \frac{\partial \sum \tilde{\mathcal{\bm{A}}}_{(c, i)}^{\text{contrastive}}}{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}} \end{bmatrix}_{i}^T \end{bmatrix}_{j,k}}_{\mathbb{R}^{|C| \times D_1 D_2}}
			\cdot \underbrace{\frac{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}}{\partial \theta}}_{\mathbb{R}^{D_1D_2 \times P}} \\
		&= - \begin{bmatrix} \frac{\text{exp}\left( \sum \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}\right)}{\sum_j \text{exp}\left({\sum \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, j)}}\right)} \end{bmatrix}_i \cdot \bm{1}_{|C| \times D_1 D_2} \cdot \frac{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}}{\partial \theta} \\
			&= - \sigma\left( \begin{bmatrix} \sum \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} \end{bmatrix}_i \right) \cdot \bm{1}_{|C| \times D_1 D_2} \cdot \frac{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}}{\partial \theta}
	\end{align}
	Each element within the second term is constant $\forall j \in \mathbb{N}(D_1),~k \in \mathbb{N}(D_2)$, indicating the gradient contribution from every given pixel $(j, k)$ from the attribution map is equivalent. This \underline{holds irrespective of whether the given pixel contains the target feature}; consequently, any region within the image can contribute towards the correct class.
\end{proof}
	Let $\gamma \in [0, 1]$ be the proportion of the attribution map containing the target. The so-called `valid' gradient contributions will only come from the regions containing the target, and smaller targets from images taken from far away result in a smaller $\gamma$, and as a consequence a greater proportion of non-target regions are established as new priors over which the classification is determined.

@@

\paragraph{Core Risk Minimization \citep{singla2022core}} proposes a generalization of Emprical Risk Minimization, the objective is to minimize prediction error while non-core regions are injected with gaussian noise:
\begin{equation}
	\min_\theta \mathbb{E}_{(\bm{x}, \bm{h}, y)} [ \mathbb{E}_{\bm{z} \sim \mathcal{N}(0, \sigma^2 \bm{I})} \ell(f_\theta(\bm{x} + \bm{z} \odot (1 - \bm{h})), y) ]
\end{equation}
Here $\ell$ represents cross-entropy, $\bm{x}$ represents the input, $\bm{h}$ is the set of core features corresponding to the input encoded as a binary mask, $y$ is the label corresponding to the input and $f_\theta$ is the parameterized network.

@@

The model should, through the process of training learn to distinguish and subsequently ignore the background in determining the final prediction. Analogous to \citep{singla2022core}, we define the core-constrained optimization objective as follows:
\begin{equation}\label{icrm}
	\min_\theta \mathbb{E}_{(\bm{x}, \bm{h}, y)} \left[ \int_{\mathbb{R}^{C \times D_1 \times D_2}} \ell(f_\theta(\bm{x} + \bm{z} \odot (1 - \bm{h})), y)~d\bm{z} \right]
\end{equation}

@@

In the following subsections, we outline cost functions designed to minimize (\ref{icrm}) using contrastive maps to directly learn invariance to spurious features while generating contrast over the target region.

\subsubsection{Divergence Minimization}
Segmentation maps describe the region over which the classification target is present. We can leverage this information to define a constrastive cost function as follows:
\begin{gather}
	\{(x, (h, y))\}^N \in \mathcal{D} \text{ where } h_{i,j} := \begin{cases}
		1 & \text{pixel contains subject} \\
		0 & \text{otherwise}
	\end{cases} \\
	D_{KL}(P || Q) := \sum_{i,j} P \cdot \log \frac{P'}{Q} \text{ where } 
	P'_{i,j} := \begin{cases}
		1 & h_{i,j} \neq 0 \\
		P_{i,j} & \text{otherwise}
	\end{cases} \label{d_kl} \\
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \frac{1}{|C|} \sum^{|C|}_{i} \underbrace{\lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_3 h)\ ||\ \sigma_{\text{softmax}}(\lambda_2 \tilde{\mathcal{\bm{A}}}^{\textit{contrastive}}_{(c, i)}))}_{\text{minimize divergence from target shape}} \\
		&+ \underbrace{\frac{\lambda_4}{(|C| - 1) \left(D_1 D_2 - \sum^{D_1,D_2}_{i,j}h_{i,j}\right)} || (1 - h) \odot \tilde{\mathcal{\bm{A}}}^{\textit{contrastive}}_{(c, i)}||^2_F}_{\text{minimize background influence}} \\
		&- \underbrace{\frac{\lambda_5}{(|C| - 1) \sum^{D_1,D_2}_{i,j}h_{i,j}} \sum^{|C|-1,D_1,D_2}_{i,j,k} \left(\tilde{\mathcal{\bm{A}}}^{\textit{contrastive}}_{(c,i)} \odot h\right)_{j,k} }_{\text{maximize foreground influence}}
	\end{split}
\end{gather}

@@

\begin{theorem}\label{ace}
	We can integrate background invariance to the definition of cross-entropy using the following formulation:
	\begin{gather}
		\mathcal{J}(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left(-\sum h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} + \sum |(1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}|\right)} \right)
	\end{gather}
\end{theorem}
\begin{proof} Following the setup from Theorem \ref{ce-spur-corr}, for a target class $c$, cross-entropy loss is defined as:
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \sigma_c(\bm{s}) \right) = -\log \left( \frac{1}{\sum_i \text{exp}({-\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}})} \right)
	\end{gather}
	We can break down each contrastive map to foreground and background components:
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} - \sum (1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}\right)} \right)
	\end{gather}
	The current loss formulation motivates improving contrast using either the foreground or the background. We penalize contrast on the background, motivating learning invariance to noise in the form of the background.
	\begin{gather}
		\mathcal{J}(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} + \sum |(1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}|\right)} \right)
	\end{gather}
	This gives us our ablated cross-entropy formulation.
\end{proof}

@@

\begin{gather}
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \frac{1}{|C|} \sum^{|C|}_{i} \lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_3 h)\ ||\ \sigma_{\text{softmax}}(\lambda_2 \tilde{\mathcal{\bm{A}}}^{\textit{contrastive}}_{(c, i)})) \\
		&- \lambda_4 \log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} + \sum |(1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}|\right)} \right)
	\end{split}
\end{gather}

@@

Similar to ACE, we define a divergence term for the target class to motivate learning the foreground as part of the training process.
\begin{gather}
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \frac{1}{|C|} \sum^{|C|}_{i} \lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_3 h)\ ||\ \sigma_{\text{softmax}}(\lambda_2 \tilde{\mathcal{\bm{A}}}_{i}^{\text{HiResCAM}})) \\
		&- \lambda_2 \bigg[ y_i \log \left(\sigma\left(h \odot \sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - (1-h) \odot \sum_{j,k}|\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}}|\right) \right) + (1 - y_i) \log \left( \bm{s}_i \right) \bigg]
	\end{split}
\end{gather}

@@

@@

\section{Results}

>>>>

\newcommand{\contcam}[2]{\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(#1, #2)}}
\newcommand{\vectorproj}[2][]{\text{proj}_{\vect{#1}}\vect{#2}}
\newcommand{\vect}{\mathbf}

@@

Per \cite{draelos2020use}, a HiResCAM is defined as follows:
\begin{equation}
	\tilde{\mathcal{\bm{A}}}_c^{\text{HiResCAM}} = \sum^F_{f=1} \frac{\partial s_c}{\partial \bm{A}^f} \odot \bm{A}^f
\end{equation}

Where $s_c$ denotes the logit for a single class. $\tilde{\mathcal{\bm{A}}}_c^{\text{HiResCAM}}$ generates a class-specific CAM. For the CAM architecture, \cite{draelos2020use} (section 3.6.2) shows that:
\begin{equation}
	\sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} = \frac{\bm{w}_c \bm{A}_F}{D_1 D_2} \\
	s_c = \frac{\bm{w}_c \bm{A}_F}{D_1 D_2} + b_c = \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} + b_c
\end{equation}
Therefore, we can recreate the entire input-dependent part of the score using HiResCAMs. Doing this for each class $c$, we obtain the bias-removed logit vector $\in \mathbb{R}^{c}$.

@@

\begin{theorem}\label{classprobdiff}
	Individual class probabilites $\sigma_c(\bm{s})$ depend on the difference(s) between class logits.
\end{theorem}
\begin{proof} Individual class probabilities for logit vector $\bm{s}$ are defined as:
	\begin{equation}
		\sigma_c(\bm{s}) = \frac{e^{s_c}}{\sum_i e^{s_i}}
	\end{equation}
	We define our logit vector in terms of the elementwise difference to a target class $c$:
	\begin{equation}
		\bm{d} = [s_i - s_c]_i \\
		\bm{s} = [s_c + d_i]_i
	\end{equation}
	Based on this definition, class probabilities can now be computed as:
	\begin{align}
		\sigma_c(\bm{s}) &= \frac{e^{s_c}}{\sum_i e^{s_c+d_i}} \\
				 &= \frac{e^{s_c}}{e^{s_c} \sum_i e^{d_i}} = \frac{1}{\sum_i e^{d_i}}
	\end{align}
	This recontextualizes softmax as a direct function of the differences of class logits.
\end{proof}
\begin{corollary} Since class probabilities depend on logit differences, the class probabilities are identitical for any two pairs of logit vectors with identical logit differences.
	\begin{equation}
		\frac{e^{s_c}}{\sum_i e^{s_i}} = \frac{e^{s_c+r}}{\sum_i e^{s_i+r}} \forall r \in \mathbb{R}
	\end{equation}
\end{corollary}
\begin{proof} Recontextualizing $\bm{s}$ per the previous proposition, we get:
	\begin{align}
		\frac{e^{s_c}}{\sum_i e^{s_c + d_i}} &= \frac{e^{s_c+r}}{\sum_i e^{s_c+d_i+r}} \\
		\frac{e^{s_c}}{e^{s_c}\sum_i e^{d_i}} &= \frac{e^{s_c+r}}{e^{s_c + r} \sum_i e^{d_i}} \\
		\frac{1}{\sum_i e^{d_i}} &= \frac{1}{\sum_i e^{d_i}}
	\end{align}
	Therefore, the absolute scale of a logit does not influence the final class probability.
\end{proof}

@@

We therefore derive an optimization objective that attempts to maximize the contrast between the two activations maps. For the binary classification case, $(c, c')$ reflects that correct and incorrect class respectively.
\begin{equation}
	\max_{\theta} s_c - s_{c'} \equiv \max_{\theta} \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} + b_c - \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c',i,j}^{\text{HiResCAM}} - b_{c'}
\end{equation}
Ignoring the bias terms, the new optimization objective becomes:
\begin{equation}
	\max_{\theta} \sum^{D_1,D_2}_{i,j} \contcam{c}{c'}{}_{, i, j} \quad \text{where} \quad \contcam{c}{c'} := \tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}}
\end{equation}
This is equivalent to the original ERM optimization objective, with the difference being the resultant values preserving spatial information in the form of the activation map instead of flattening the feature map. If $\tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}} > 0$, the classification is correct -- if not, the classification is incorrect.

@@

\begin{definition}
	Given a set of classes $C$ with $c_t$ being the target class and $c_{t'}$ being a sampled non-target class, we define ContrastiveCAMs as follows:
	\begin{equation}
		\contcam{c_t}{c_{t'}} := \left\{\tilde{\mathcal{\bm{A}}}_{c_t}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c_{t'}}^{\text{HiResCAM}}\right\}^{|C|-1}_{c_{t'} \in c \setminus c_t}
	\end{equation}
\end{definition}

@@

To ensure that the model accurately predicts the target class, each contrastive comparison should yield a positive result. Therefore, the multiclass optimization objective under this construction reads:
\begin{equation}
	\max_{\theta} \sum^{D_1,D_2}_{j,k}\tilde{\mathcal{\bm{A}}}_{(c, c'),j,k}^{\text{contrastive}}\ \forall c' \in \mathbb{Z}_{+}(|C| - 1)
\end{equation}

@@

\begin{theorem}\label{ce-spur-corr}
	Cross-Entropy motivates learning from spurious correlations.
\end{theorem}
\begin{proof} Using the definition of HiResCAMs, ContrastiveCAMs and the elementwise difference setup from Theorem \ref{classprobdiff}, we have:
	\begin{align}
		d &= [s_i - s_c]_i \\
		&= \left[\sum^{D_1,D_2}_{j,k} \tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - \sum^{D_1,D_2}_{j,k} \tilde{\mathcal{\bm{A}}}_{c,j,k}^{\text{HiResCAM}}\right]_i \\
		&= \left[ -\sum \contcam{c}{i} \right]_i
	\end{align}
	As an aside: we note that when $i = c$, the contrastive map obtained is $\bm{0} \in \mathbb{R}^{D_1 \times D_2}$, and instead of $|C| - 1$ contrastive maps we now obtain $|C|$ contrastive maps.

	Using our result from Theorem~\ref{classprobdiff}, we have:
	\begin{equation}
		\sigma_c(\bm{s}) = \frac{1}{\sum_i e^{d_i}} = \frac{1}{\sum_i \text{exp}({-\sum \contcam{c}{i}})} = \sum_i \text{exp}\left({\sum \contcam{c}{i}}\right)
	\end{equation}
	For a target class $c$, cross-entropy loss is defined as:
	\begin{equation}
		H(\theta, \mathcal{D}) = -\log \left( \sigma_c(\bm{s}) \right) = -\log \left( \sum_i \text{exp}\left({\sum \contcam{c}{i}}\right) \right)
	\end{equation}
	Vectorizing parameters $\theta \in \mathbb{R}^P$ as well as the contrastive attribution maps $\tilde{\mathcal{\bm{A}}}_{(c, i)}^{\text{contrastive}} \in \mathbb{R}^{D_1 D_2}$, we can use the above result to compute gradient contributions from each pixel within the attribution map:
	\begin{align}
		\nabla_\theta H &= \underbrace{\begin{bmatrix} \frac{\partial H}{\partial \sum \tilde{\mathcal{\bm{A}}}_{(c, i)}^{\text{contrastive}}} \end{bmatrix}_i}_{\mathbb{R}^{1 \times |C|}}
			\cdot \underbrace{\begin{bmatrix} \begin{bmatrix} \frac{\partial \sum \tilde{\mathcal{\bm{A}}}_{(c, i)}^{\text{contrastive}}}{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}} \end{bmatrix}_{i}^T \end{bmatrix}_{j,k}}_{\mathbb{R}^{|C| \times D_1 D_2}}
			\cdot \underbrace{\frac{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}}{\partial \theta}}_{\mathbb{R}^{D_1D_2 \times P}} \\
		&= - \begin{bmatrix} \frac{\text{exp}\left( \sum \contcam{c}{i}\right)}{\sum_j \text{exp}\left({\sum \contcam{c}{j}}\right)} \end{bmatrix}_i \cdot \bm{1}_{|C| \times D_1 D_2} \cdot \frac{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}}{\partial \theta} \\
			&= - \sigma\left( \begin{bmatrix} \sum \contcam{c}{i} \end{bmatrix}_i \right) \cdot \bm{1}_{|C| \times D_1 D_2} \cdot \frac{\partial \tilde{\mathcal{\bm{A}}}_{(c, i), j, k}^{\text{contrastive}}}{\partial \theta}
	\end{align}
	Each element within the second term is constant $\forall j \in \mathbb{N}(D_1),~k \in \mathbb{N}(D_2)$, indicating the gradient contribution from every given pixel $(j, k)$ from the attribution map is equivalent. This \underline{holds irrespective of whether the given pixel contains the target feature}; consequently, any region within the image can contribute towards the correct class.
\end{proof}
	Let $\gamma \in [0, 1]$ be the proportion of the attribution map containing the target. The so-called `valid' gradient contributions will only come from the regions containing the target, and smaller targets from images taken from far away result in a smaller $\gamma$, and as a consequence a greater proportion of non-target regions are established as new priors over which the classification is determined.

@@

\paragraph{Core Risk Minimization \citep{singla2022core}} proposes a generalization of Emprical Risk Minimization, the objective is to minimize prediction error while non-core regions are injected with gaussian noise:
\begin{equation}
	\min_\theta \mathbb{E}_{(\bm{x}, \bm{h}, y) \sim \mathcal{D}} [ \mathbb{E}_{\bm{z} \sim \mathcal{N}(0, \sigma^2 \bm{I})} \ell(f_\theta(\bm{x} + \bm{z} \odot (1 - \bm{h})), y) ]
\end{equation}
Here $\ell$ represents cross-entropy, $\bm{x}$ represents the input, $\bm{h}$ is the set of core features corresponding to the input encoded as a binary mask, $y$ is the label corresponding to the input and $f_\theta$ is the parameterized network.

@@

The model should, through the process of training learn to distinguish and subsequently ignore the background in determining the final prediction. Analogous to \citep{singla2022core}, we define the contrastive-constrained optimization objective as follows:
\begin{equation}\label{ccrm}
	\min_\theta \mathbb{E}_{(\bm{x}, \bm{h}, y) \sim \mathcal{D}} \left[ \ell(f_\theta(\bm{x}, y) \right] \quad \text{s.t.} \quad \sum_{c' \in C} \| \contcam{c}{c'} \odot (1 - h) \| = 0
\end{equation}

@@

In the following subsections, we outline a family of cost functions for binary, mutliclass and mutliple-class classification, designed to minimize (\ref{ccrm}) using contrastive maps to learn invariance to spurious features while generating contrast over the target region.

\subsubsection{Divergence Minimization}
Segmentation maps describe the region over which the classification target is present. We can leverage this information to define a constrastive cost function as follows:
\begin{gather}
	\{(x, (h, y))\}^N \in \mathcal{D} \text{ where } h_{i,j} := \begin{cases}
		1 & \text{pixel contains subject} \\
		0 & \text{otherwise}
	\end{cases} \\
	D_{KL}(P || Q) := \sum_{i,j} P \cdot \log \frac{P'}{Q} \text{ where } 
	P'_{i,j} := \begin{cases}
		1 & h_{i,j} \neq 0 \\
		P_{i,j} & \text{otherwise}
	\end{cases} \label{d_kl} \\
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \frac{1}{|C|} \sum^{|C|}_{i} \underbrace{\lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_3 h)\ ||\ \sigma_{\text{softmax}}(\lambda_2 \contcam{c}{i}))}_{\text{minimize divergence from target shape}} \\
		&+ \underbrace{\frac{\lambda_4}{(|C| - 1) \left(D_1 D_2 - \sum^{D_1,D_2}_{i,j}h_{i,j}\right)} || (1 - h) \odot \contcam{c}{i}||^2_F}_{\text{minimize background influence}} \\
		&- \underbrace{\frac{\lambda_5}{(|C| - 1) \sum^{D_1,D_2}_{i,j}h_{i,j}} \sum^{|C|-1,D_1,D_2}_{i,j,k} \left(\contcam{c}{i} \odot h\right)_{j,k} }_{\text{maximize foreground influence}}
	\end{split}
\end{gather}

@@

\begin{theorem}\label{ace}
	We can integrate background invariance to the definition of cross-entropy using the following formulation:
	\begin{equation}
		\mathcal{J}(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left(-\sum h \odot \contcam{c}{i} + \sum |(1-h) \odot \contcam{c}{i}|\right)} \right)
	\end{equation}
\end{theorem}
\begin{proof} Following the setup from Theorem \ref{ce-spur-corr}, for a target class $c$, cross-entropy loss is defined as:
	\begin{equation}
		H(\theta, \mathcal{D}) = -\log \left( \sigma_c(\bm{s}) \right) = -\log \left( \frac{1}{\sum_i \text{exp}({-\sum \contcam{c}{i}})} \right)
	\end{equation}
	We can break down each contrastive map to foreground and background components:
	\begin{equation}
		H(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum h \odot \contcam{c}{i} - \sum (1-h) \odot \contcam{c}{i}}\right)} \right)
	\end{equation}
	The current loss formulation motivates improving contrast using either the foreground or the background. We penalize contrast on the background, motivating learning invariance to noise in the form of the background.
	\begin{equation}
		\mathcal{J}(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum h \odot \contcam{c}{i} + \sum |(1-h) \odot \contcam{c}{i}}|\right)} \right)
	\end{equation}
	This gives us our ablated cross-entropy formulation.
\end{proof}

@@

\begin{equation}
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \frac{1}{|C|} \sum^{|C|}_{i} \lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_3 h)\ ||\ \sigma_{\text{softmax}}(\lambda_2 \contcam{c}{i})) \\
		&- \lambda_4 \log \left( \frac{1}{\sum_i \text{exp}\left({-\sum h \odot \contcam{c}{i} + \sum |(1-h) \odot \contcam{c}{i}}|\right)} \right)
	\end{split}
\end{equation}

@@

Similar to ACE, we define a divergence term for the target class to motivate learning the foreground as part of the training process.
\begin{equation}
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \frac{1}{|C|} \sum^{|C|}_{i} \lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_3 h)\ ||\ \sigma_{\text{softmax}}(\lambda_2 \tilde{\mathcal{\bm{A}}}_{i}^{\text{HiResCAM}})) \\
		&- \lambda_2 \bigg[ y_i \log \left(\sigma\left(h \odot \sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - (1-h) \odot \sum_{j,k}|\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}}|\right) \right) + (1 - y_i) \log \left( \bm{s}_i \right) \bigg]
	\end{split}
\end{equation}

@@

The application of Optimal Transport theory as a robust metric between distributions enables vast applications within image processing \citep{papadakis2015optimal}. Robust gradients in turn improve convergence, which we leverage by adapting the spatial sliced wasserstein loss proposed in \citet{heitz2021sliced} to learn the optimal contrastive map as a regularizer term over (\ref{ace}).
\begin{equation}
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \mathbb{E}_{\mathbf{d} \sim \mathcal{N}(0, 1)}\left[\left\| \text{sort}(\vectorproj[d]{(\text{concat}(\contcam{c}{c'}, \bm{s}}))) - \text{sort}(\vectorproj[d]{\text{concat}(h, s)}) \right\|^2 \right] \\
		&-\log \left( \frac{1}{\sum_i \text{exp}\left(-\sum h \odot \contcam{c}{i} + \sum |(1-h) \odot \contcam{c}{i}|\right)} \right)
	\end{split}
\end{equation}
Where $\mathbf{s}$ is a spatial prior that ensures learned distributions correspond to spatial regions that contain the target. In practice, we use sample $\approx 20$ directions, and exclude over-activation in target regions when computing the norm.

We observe that using divergence / distance based regularization to learn segmentation maps as an implicit part of optimization improves the generalization performance of the network. We hypothesize this being a consequence of inducing contrast over regions beyond just those most relevant for classification within dataset-specific samples. We present empricial evidence to support this claim across all our benchmarks in Section \ref{results}.

@@

\section{Results}
\label{results}
